01_pca

### 2019-04-23 analysis/01_pca_all.Rmd
add pca figs

```{r setup}
library(ggfortify) # for PCA analysis
library(cluster)

@@

```{r rawdata}
colData <- read.csv("../results/00_colData_characterization.csv", header = T, row.names = 1)
countData <- read.csv("../results/00_countData_characterization.csv", header = T, row.names = 1)
head(colData)
geneinfo <- read.csv("../results/00_geneinfo.csv", row.names = 1)
```

@@

# pca
pca <- prcomp(t(countData))
autoplot(pca, data = colData, colour = 'tissue') 
autoplot(pca, data = colData, colour = 'sex') 
autoplot(pca, data = colData, colour = 'treatment') 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex") 
autoplot(pca, data = colData, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = colData, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(countData),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(countData), 3), frame = TRUE, frame.type = 'norm')
```

>>>>

```{r setup}
library(tidyverse)
library(ggfortify) # for PCA analysis
library(cluster)
library(cowplot)

theme_rmh <- function(){ 
    theme_bw(base_size=10) +
        theme(
            panel.grid.minor.x  = element_blank(),
           panel.grid.minor.y  = element_blank(),
            strip.background = element_rect(colour="white", fill="white"),
              legend.position = "right",           
           legend.margin=margin(t=-0.1, r=0, b=-0.1, l=-0.1, unit="cm"),
           legend.key.size = unit(0.5, "cm"))
}

@@

```{r rawdata}
counts <- read.csv("../results/00_countData_characterization.csv", row.names = 1, header = T)

samples <- read.csv("../metadata/00_colData_characterization.csv", row.names = 1, header = T)
geneinfo <- read.csv("../metadata//00_geneinfo.csv", row.names = 1, header = T)

head(samples)
```

@@

# pca
pca <- prcomp(t(counts))
autoplot(pca, data = samples, colour = 'tissue') 
autoplot(pca, data = samples, colour = 'sex') 
autoplot(pca, data = samples, colour = 'treatment') 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex") 
autoplot(pca, data = samples, colour = 'tissue', shape = "sex", alpha = "treatment")
autoplot(pca, data = samples, colour = 'treatment', shape = "sex", alpha = "tissue")

# kmeans
autoplot(clara(t(counts),3), frame = TRUE)

# probability ellipse
autoplot(pam(t(counts), 3), frame = TRUE, frame.type = 'norm')
```

```{r pca-es}
p <- autoplot(pca, data = samples, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = "PCA plot in English") +
  scale_colour_discrete(name = "colour", labels = c("rosado", "verde", "azul")) +
  scale_shape_discrete(name = "shape", labels = c("circle", "square")) +
  theme_rmh()

es <- p + labs(subtitle = "PCA plot en español") +
  scale_colour_discrete(name = "color", labels = c("rosado", "verde", "azul")) +
  scale_shape_discrete(name = "forma", labels = c("circulo", "triangulo")) +
  theme_rmh()

enes <- plot_grid(en,es)

enes

ggsave("../figures/espanol/pca-en-es.png", width=6, height=3, dpi=300)

pdf(file="../figures/espanol/pca-en-es.pdf", width=6, height=3)
plot(enes)    
dev.off()
```

```{r subset}
for (eachgroup in levels(samples$treatment)){

  print(eachgroup)

  colData <- samples %>%
      dplyr::filter(treatment == eachgroup) %>%
      droplevels()
  row.names(colData) <- colData$V1

  savecols <- as.character(colData$V1) 
  savecols <- as.vector(savecols)

  countData <- counts %>% dplyr::select(one_of(savecols))

  # check that row and col lenghts are equal
  print(ncol(countData) == nrow(colData))

  pca <- prcomp(t(countData))

  p <- autoplot(pca, data = colData, colour = 'tissue', shape = "sex")

en <- p + labs(subtitle = eachgroup) +
  theme_rmh()

plot(en) 
}
```
